---
title: "It's About Time"
format:
  html:
    code-fold: true
    code-line-numbers: true
editor: visual
execute: 
  keep-md: true
warnings: false
---

```{r}
library(tidyverse)
library(lubridate)

sales <- read.csv("https://byuistats.github.io/M335/data/sales.csv")

sales2 <- sales %>% 
  filter(Name != "Missing") %>% 
  filter(Time != "2016-04-20T19:01:00Z") %>% 
  filter(Time != "2016-05-13T20:27:00Z") %>% 
  mutate(sale_time = force_tz(ymd_hms(Time), tzone = "America/Denver")) %>% 
  filter(Amount >0) %>% 
  rename("Company" = "Name") %>% 
  separate(sale_time, into = c("date", "timestamp"), sep = -8, remove = FALSE)
  
hourly_sales <- sales2 %>% 
  mutate(hourly = hour(sale_time))

transactions <- sales2 %>% 
  group_by(Company) %>% 
  summarise(Sales = n())


#group_by name and date

```

## Gross Revenue Over Time per Company

```{r}
ggplot(data = sales2, aes(x = sale_time, y = Amount, group = Company, color = Company)) +
  geom_line(size = .75) +
  scale_y_continuous(trans = "sqrt") +
  labs(title = "Gross Revenue Per Company over Three Months",
       x = NULL, 
       y = "Gross Revenue (US $)") +
  theme_minimal() +
  theme(legend.position = c(0.2, .78))

```

## Transactions per Company


```{r}
ggplot(data = transactions, aes(x = Company, y = Sales, fill = Company)) + 
  geom_col() +
  labs(title = "Transactions per Company", 
       y = "Transactions") +
  theme_minimal()


```

## Hours of Operation

Hours of Operation

```{r}
ggplot(hourly_sales, aes(x = hourly, y = Amount, color = Company, groups = Company)) +
  geom_point() +
  labs(title = "Hours of Operation per Company", 
       x = "Time of Day",
       y = "Transaction Amount (US $)") +
  scale_y_continuous(trans = "sqrt",
                     limits = c(0, 800)) +
  facet_wrap(~Company, nrow = 2, ncol = 3) +
  theme_minimal()

#scale_x_continuous(breaks = seq(0, 24, 2)) +

```


I would give a loan to LeBelle. They seemed to perform the best because they have fewer transactions, but more income than the other companies at a higher rate. This tells me that their products are more expensive, but they seem to have a consistent rate of sales over the three months. At the very end of the three month period, they also experienced a sharp increase of sales, reaching over $900 in revenue in just one day. In addition, they seem to have a steady amount of transactions throughout all hours of operation.