---
title: "US Cities"
format: 
  html:
    code-line-numbers: true
    code-fold: true
editor: visual
execute: 
  keep-md: true
  keep-png: true
warnings: false
---

```{r}
library(tidyverse)
library(sf)
library(USAboundaries)
library(ggrepel)

```

```{r}
idaho <- us_counties(states = "Idaho")


US <- us_states() %>% 
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii", "District of Columbia"))

US_cities <- us_cities() %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii", "District of Columbia")) %>% 
  select(c(city, state_name, population)) %>% 
  group_by(state_name) %>% 
  slice_max(population, n = 3) %>% 
  arrange(desc(population), .by_group = TRUE) %>% 
  mutate(rank = c(1, 2, 3))

city_labels <- US_cities %>% 
  slice_max(population, n = 1)

# assignment: color = rank, names of cities = ggrepel and biggest city labelled

```


```{r}

ggplot() + 
  geom_sf(data = US, fill = NA, color= "gray35") +
  geom_sf(data = idaho, fill = NA, color = "gray35") +
  geom_sf(data = US_cities, aes(color = rank, size = population/1000)) +
  geom_label_repel(data = city_labels, aes(label = city, geometry = geometry),
                   stat = "sf_coordinates", 
                   size = 1, 
                   color = "darkblue",
                   force = 1,
                   force_pull = 0.3,
                   max.overlaps = Inf) +
  labs(x = NULL,
       y = NULL,
       color = NULL,
       size = "Population (1,000)") +
  guides(color = "none",
         size = guide_legend(override.aes = list(color = "darkblue"))) +
  theme_bw()
ggsave(filename = "US-Cities-Map.png", width = 15, height = 10)

```

I chose to use ChatGPT AI to assist me with this assignment by tutoring me on how to create a basic map chart and finding ways to use code when the textbook didnt have examples to guide me. I chose to use this tool because it could give me examples or help when I couldn't find something in the book, and it affected my learning by teaching me how to use new functions I hadn't tried before for a new type of assignment.

