---
title: "Recreate - Visualization"
format: 
  html:
    code-line-numbers: true
    code-fold: true
execute: 
  keep-md: true
editor: visual
warnings: false
---

```{r}
library(tidyverse)
library(countrycode)
library(ggplot2)
library(ggrepel)
library(gganimate)
library(scales)

```

```{r}
df_fruit <- read_csv("https://ourworldindata.org/grapher/fruit-consumption-vs-gdp-per-capita.csv?v=1&csvType=full&useColumnShortNames=true")
```

```{r}
fruit <- df_fruit %>%
  mutate(Continent = countrycode(Entity, origin = "country.name", destination = "continent")) %>% 
  filter(!is.na(ny_gdp_pcap_pp_kd)) %>%
  filter(!is.na(Continent)) %>% 
  rename(fruit_supply = `fruit__00002919__food_available_for_consumption__0645pc__kilograms_per_year_per_capita`) %>%
  filter(!is.na(fruit_supply)) %>% 
  filter(Year == 2022)

```

## Fruit consumption vs. GDP per capita

```{r}
ggplot(fruit, aes(x = ny_gdp_pcap_pp_kd, y = fruit_supply, color = Continent)) +
  geom_point() +
  scale_y_continuous(breaks = seq(0, 500, 100),
                     limits = c(0, 500),
                     labels = label_number(suffix = " kg")) +
  scale_x_continuous(breaks = c(1000, 2000, 5000, 10000, 20000, 50000, 100000),
                     labels = dollar_format(prefix = "$", big.mark = ",", accuracy = 1),
                     trans = "log10") +
  labs(title = "Fruit consumption vs. GDP per capita",
       subtitle = "Average per capita fruit supply, measured in kilograms per year versus gross domestic product (GDP) per capita, adjusted \nfor inflation and for differences in living costs between countries.",
       x = "GDP per capita (international-$ in 2021 prices)", 
       y = "Fruit supply per person (kilograms per year)",
       caption = "Data source: Food and Agriculture Organization of the United Nations (2024); Data compiled from multiple sources by World Bank (2025) – Learn more about this data \nNote: GDP per capita is expressed in international-$ at 2021 prices.
\nOurWorldinData.org/diet-compositions | CC BY") +
  coord_cartesian(xlim = c(1000, 100000)) +
  theme_minimal()
```

```{r}
specific <- fruit %>% 
  filter(Entity %in% c("Kenya", "Ghana", "United States"))

ggplot(fruit, aes(x = ny_gdp_pcap_pp_kd, y = fruit_supply, color = Continent)) +
  geom_point() +
  geom_point(data = specific, aes(x = ny_gdp_pcap_pp_kd, y = fruit_supply), pch = 21, size = 2, color = "black") +
  scale_y_continuous(breaks = seq(0, 500, 100),
                     limits = c(0, 500),
                     labels = label_number(suffix = " kg")) +
  scale_x_continuous(breaks = c(1000, 2000, 5000, 10000, 20000, 50000, 100000),
                     labels = dollar_format(prefix = "$", big.mark = ",", accuracy = 1),
                     trans = "log10") +
  labs(title = "Fruit consumption vs. GDP per capita: Kenya, Ghana, United States",
       subtitle = "Average per capita fruit supply, measured in kilograms per year versus gross domestic product (GDP) per capita, adjusted \nfor inflation and for differences in living costs between countries.",
       x = "GDP per capita (international-$ in 2021 prices)", 
       y = "Fruit supply per person (kilograms per year)",
       caption = "Data source: Food and Agriculture Organization of the United Nations (2024); Data compiled from multiple sources by World Bank (2025) – Learn more about this data \nNote: GDP per capita is expressed in international-$ at 2021 prices.
\nOurWorldinData.org/diet-compositions | CC BY") +
  coord_cartesian(xlim = c(1000, 100000)) +
  theme_minimal()

```

## Oceania

```{r}
specific2 <- fruit %>% 
  filter(Continent == "Oceania")

ggplot(fruit, aes(x = ny_gdp_pcap_pp_kd, y = fruit_supply, color = Continent)) +
  geom_point() +
  geom_point(data = specific2, aes(x = ny_gdp_pcap_pp_kd, y = fruit_supply), pch = 21, size = 2, color = "black") +
  scale_y_continuous(breaks = seq(0, 500, 100),
                     limits = c(0, 500),
                     labels = label_number(suffix = " kg")) +
  scale_x_continuous(breaks = c(1000, 2000, 5000, 10000, 20000, 50000, 100000),
                     labels = dollar_format(prefix = "$", big.mark = ",", accuracy = 1),
                     trans = "log10") +
  labs(title = "Fruit consumption vs. GDP per capita: Oceania",
       subtitle = "Average per capita fruit supply, measured in kilograms per year versus gross domestic product (GDP) per capita, adjusted \nfor inflation and for differences in living costs between countries.",
       x = "GDP per capita (international-$ in 2021 prices)", 
       y = "Fruit supply per person (kilograms per year)",
       caption = "Data source: Food and Agriculture Organization of the United Nations (2024); Data compiled from multiple sources by World Bank (2025) – Learn more about this data \nNote: GDP per capita is expressed in international-$ at 2021 prices.
\nOurWorldinData.org/diet-compositions | CC BY") +
  coord_cartesian(xlim = c(1000, 100000)) +
  theme_minimal()


```

This was a challenging assignment, and I still feel that I could've done better with more time. I discovered that the chart looks very different when using a linear vs. logarithmic scale, because the website default is logarithmic but the linear model matches mine much closer. I also realized I am not confident in using axis, plot, and legend aspects to polish a graph. However, I did learn how to create a moving graph that tells the story over time, which I ended up deciding to remove, and this was an interesting tool to try out. Emphasizing different aspects of data also draws the eyes and highlights patterns and gives more meaning to the data. I used layering to highlight certain points, as well as size adjustments and color contrast.
